
---
title: "Heart Disease Prediction"
author: "Anna Wallin"
date: "Jan 2021"
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---

```{r install, message = FALSE}

library(readr)
```

```{r}
mydata<-read_csv("HeartDiseaseOG.csv") #load in data
#binary response variable is TenYearCHD
#Predictor variables are: male, age, education, currentSmoker,cigsPerDay,BPMeds, prevalentStroke,prevalentHyp,diabetes, totChol, sysBP, diaBP, BMI, heartRate, glucose
#education has a value of 1 to 4 depending on if the respondent did some high school, high school or GED, some college or vocational school, or college

summary(mydata) #summary statistics of the data set
sapply(mydata,sd) #view standard deviations
```

```{r logit}
mydata$education<-factor(mydata$education) #make education categorical var 

#run logistic regression
logit_model<-glm(TenYearCHD~male+age+education+currentSmoker+cigsPerDay+BPMeds+ prevalentStroke+prevalentHyp+diabetes+totChol+sysBP+diaBP+BMI+heartRate+glucose, data=mydata,family = binomial(link = "logit"))


```

```{r view}
summary(logit_model)

```
# Variables that are statistically significant
Variables that have a p-value below 0.05: male, age, cigsPerDay, totChol, sysBP, and glucose.

The coefficients show the log odds for a one unit increase in the x (predictor) variable.

For example, for a one unit increase in cigsPerDay, being categorized as at risk for coronary heart disease over a ten year period increases by 0.018387.


```{r CI}
#calc CIs
confint(logit_model)

```


```{r CI default}

confint.default(logit_model)

```

```{r odds ratio }
exp(coef(logit_model))

```
```{r odds ratio table with CI}
exp(cbind(Odds_Ratio=coef(logit_model),confint(logit_model))) # 95 % CI

```
You can exponentiate coefficients to find odds ratios. Here is the formula simplified...
\begin{displaymath}\frac{exp(\alpha+ \beta(x_i + 1))}{exp(\alpha+\beta x_i)}\end{displaymath}

The above becomes...
\begin{displaymath}=\frac{exp(\alpha+\beta x_i +\beta)}{exp(\alpha+\beta x_i)}\end{displaymath}

which further simplifies to...\
\begin{displaymath}=\frac{exp(\alpha+\beta x_i)*exp(\beta)}{exp(\alpha+\beta x_i)}\end{displaymath}

Finally...\
\begin{displaymath}=exp(\beta)\end{displaymath}

```{r exp}

exp(0.002377) #totChol odds ratio exponentiated beta
#a one unit increase in total cholesterol increases odds of CHD risk over a ten year period by 0.2377%
```

Being male increases the odds of being at risk for coronary heart disease by a factor of 1.7430891161.

A one unit increase in total cholesterol increases odds of CHD risk over a ten year period by 0.2377%


### Let's look at variable importance...

```{r var import}
#library(caret)
#update.packages()
#install.packages("caret",
                 #repos = "http://cran.r-project.org", 
                 #dependencies = c("Depends", "Imports", "Suggests"))
caret::varImp(logit_model)

```
The most important predictor of CHD is age, followed by the sex of a person.


### Now, let's examine multicollinearity
```{r mc}
#install.packages("car")

library(car)
car::vif(logit_model)
```
Looking at the Variance Inflation Indicator, all values range from 1 to 3.6. This means that there is moderate correlation between a predictor variable and other variables. Given that there is not strong correlation, this does not need any action to correct.

### Using our model to make predictions...

```{r predicted probability}

mydata_NA=na.omit(mydata)
#mean(mydata_NA$age)
#mean(mydata_NA$cigsPerDay)
#mean(mydata_NA$totChol)
#mean(mydata_NA$sysBP)
#mean(mydata_NA$glucose)
for_pred<- data.frame(age=mean(mydata_NA$age), cigsPerDay=mean(mydata_NA$cigsPerDay),totChol=mean(mydata_NA$totChol), sysBP=mean(mydata_NA$sysBP),male=1,education=factor(1), currentSmoker=1,glucose=101, BPMeds=0,prevalentHyp=0,prevalentStroke=0,diabetes=0,heartRate=0,BMI=0,diaBP=0)
predict(logit_model,for_pred,type="response")

# given average age, cigarette consumption, cholesterol, systolic blood pressure, a smoking man with above average glucose has a .24 chance of developing risk of CHD


```
these were sig: male, age, cigsPerDay, totChol, sysBP, and glucose



